---
title: "Hills Forest ReMapping"
author: "Summit-GIS"
date: "17/08/2023"
output: 
  github_document:
    keep_md: TRUE
    toc: TRUE
    toc_depth: 5
    number_sections: FALSE
    df_print: tibble
    latex_engine: xelatex
---

```{r setup, echo=FALSE, message=FALSE,warning=FALSE, error=FALSE}
library(sf)
library(sp)
library(terra)
library(raster)
library(dplyr)
library(caret)
library(caretEnsemble)
library(ForestTools)
library(lidR)
library(randomForest)
library(e1071)
library(rgdal)
library(rgeos)
library(Rcpp)
library(rmarkdown)
library(knitr)
library(MASS)
library(car)
#devtools::install_github(("gearslaboratory/gdalUtils"))
library(gdalUtils)
#library(gdalUtilities)
#webshot::install_phantomjs(force = TRUE)
#knit_hooks$set(webgl = hook_webgl)
#knit_hooks$set(rgl.static = hook_rgl)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, error=FALSE, message = FALSE)
set.seed(123)
# ShortCUt l.95
```

## Action:

Some notes on the data processing steps and mapping outputs completed in the remapping exercise of the Hills Forest Area. To map potential differences between official data layers and newly LiDAR-derived attributes, two variables were derived: 'Stocking Density (stem/ha)' & 'Age Class'. Using DEM and DSM datasets sourced from the LINZ website, a stem map was derived by applying a variable window algorithm and a default taper function. With the stem map feature, two rasters were calculated showing a canopy height model and stocking density per hectare. This allowed comparison with values at StandID level. 


## 1. Inputs: LiDAR Projection 

```{r, fig.show='hold', out.width="50%", eval=FALSE, echo=TRUE}
# Merge chunks
filez_dem = list.files("~/Desktop/Summit_Forestry/dem", full.names = T, all.files = FALSE, pattern = '.tif$') 
filez_dsm = list.files("~/Desktop/Summit_Forestry/dsm", full.names = T, all.files = FALSE, pattern = '.tif$') 
dem_raster_list <- lapply(filez_dem, raster)
dsm_raster_list <- lapply(filez_dsm, raster)
dem_raster = do.call(merge, c(dem_raster_list, tolerance = 1))
dsm_raster = do.call(merge, c(dsm_raster_list, tolerance = 1))
writeRaster(dem_raster, filename = "~/Desktop/Summit_Forestry/dem/dem_raster.tif", overwrite=TRUE)
writeRaster(dsm_raster, filename = "~/Desktop/Summit_Forestry/dsm/dsm_raster.tif", overwrite=TRUE)
dsm_rast = terra::rast(dsm_raster)
dem_rast = terra::rast(dem_raster)
elev_rast = terra::rast(dem_raster)
chm_rast = (dsm_rast - dem_rast)

# If possible, best reproject everything to DEM/DSM coordinates
# but easier to assign official wkt starting off to avoid bugs
terra::crs(elev_rast) =  "epsg:2193"
terra::crs(chm_rast) =  "epsg:2193"
terra::plot(elev_rast, main='DEM(m) - LINZ') 
terra::plot(chm_rast, main='CHM(m)') 
```

## Input: Derive AOi & Apply Mask

```{r, fig.show='hold', out.width="50%", eval=FALSE, echo=TRUE}
mask_sf = sf::read_sf("~/Desktop/Summit_Forestry/stands/HILL-0341-009.shp")
mask_rast = rasterize(vect(mask_sf), chm_rast, touches = TRUE)
mask_rast = terra::resample(mask_rast, chm_rast, method="near") # choose from one of these two
mask_raster = raster::raster(mask_rast)
writeRaster(mask_raster, filename = "~/Desktop/Summit_Forestry/stands/mask_raster_009.tif", overwrite=TRUE)
terra::crs(mask_rast) =  "epsg:2193"
# elev_rast = terra::resample(elev_rast, chm_rast, method="near")
# chm_rast = terra::resample(chm_rast, mask_rast, method="near")
ggplot(mask_sf) + geom_sf(aes(fill = 'red'), show.legend = FALSE)

# Apply Mask
elev = mask(elev_rast, mask_rast, inverse=TRUE)
chm = mask(chm_rast, mask_rast, inverse=TRUE)
elev_raster = raster::raster(elev)
chm_raster = raster::raster(chm)
writeRaster(elev_raster, filename = "~/Desktop/Summit_Forestry/dem/elev_raster.tif", overwrite=TRUE)
writeRaster(chm_raster, filename = "~/Desktop/Summit_Forestry/dsm/chm_raster.tif", overwrite=TRUE)

elev_raster = raster::raster("~/Desktop/Summit_Forestry/dem/elev_raster.tif")
chm_raster = raster::raster("~/Desktop/Summit_Forestry/dem/chm_raster.tif")

# Stack rasters & visualize
## covs_m1 = raster::stack(
##  chm_raster,
##  elev_raster, 
##  mask_raster)
## rasterVis::levelplot(covs_m1)
```

## Input: Derive Variable Window Function

```{r, fig.show='hold', out.width="50%", eval=FALSE}
# Applied Plowright's window function https://cran.r-project.org/web/packages/ForestTools/vignettes/treetop_analysis.html
# Needs function that is closer to site location/species
wf_plowright<-function(x){ 
  a=0.05
  b=0.6 
  y<-a*x+b 
  return(y)}
heights <- seq(0,40,0.5)
window_plowright <- wf_plowright(heights)
plot(heights, window_plowright, type = "l", ylim = c(0,12), xlab="point elevation (m)", ylab="window diameter (m)", main='Plowright, 2018; y=0.05*x+0.6')
```

## Input: Derive 95% Canopy Height & Stem Count Layers

```{r, eval=TRUE, fig.show='hold', out.width="33%", echo=FALSE}
kernel <- matrix(1,3,3)
chm_raster = focal(chm_rast, w = kernel, fun = median, na.rm = TRUE) %>% raster()
ttops_1.5mfloor_plowright = ForestTools::vwf(chm_raster, wf_plowright, 1.5)
writeOGR(ttops_1.5mfloor_plowright, "~/Desktop/Summit_Forestry/stands", "treetops_hills_009", driver = "ESRI Shapefile") 

quant95 <- function(x, ...) 
  quantile(x, c(0.95), na.rm = TRUE)
custFuns <- list(quant95, max)
names(custFuns) <- c("95thQuantile", "Max")

ttops_1.5mfloor_height <- ForestTools::sp_summarise(ttops_1.5mfloor_plowright, grid = 1, variables = "height", statFuns = custFuns)
chm_95height_raster = ttops_1.5mfloor_height[["height95thQuantile"]]
#chm_95height_rast = terra::rast(chm_95height_raster)
raster::writeRaster(chm_95height_raster, filename = "~/Desktop/Summit_Forestry/stands/chm_95height.tif", overwrite=TRUE)

stem_count_raster = ttops_1.5mfloor_height[["TreeCount"]]
stem_count_sf = st_as_sf(stem_count_raster)
stem_count_rast = rasterize(vect(stem_count_sf), chm_raster, fun = sum, touches = TRUE)
stem_count_ha_rast = stem_count_rast*100 #aggregate to 100m resolution for stems per hectare
stem_count_ha_raster = raster::raster(stem_count_ha_rast)
raster::writeRaster(stem_count_ha_raster, filename = "~/Desktop/Summit_Forestry/stands/stem_count_ha.tif", overwrite=TRUE)

mypalette<-brewer.pal(8,"Greens")
plot(chm_95height_raster, col = mypalette, alpha=0.6)
plot(st_geometry(stem_count_sf["treeID"]), add=TRUE, cex = 0.001, pch=19, col = 'red', alpha=0.8) 

# Point Cloud Pipeline (or smaller areas with lighter memory tasks...)
## stem_count_sp = find_trees(chm_95height, lmf(wf_plowright), uniqueness = "bitmerge")
## stem_count_sf = st_as_sf(stem_count_sp)
## mypalette<-brewer.pal(8,"Greens")
## plot(chm_95height_raster, col = mypalette, alpha=0.6)
## plot(st_geometry(stem_count_sf["treeID"]), add=TRUE, cex = 0.001, pch=19, col = 'red', alpha=0.8)

# Point Cloud Pipeline: Check Tree-ID as proof of stem count...
## algo = dalponte2016(chm_95height_raster, stem_count_sp)
## chm_95h_segmented = segment_trees(chm_95height_raster, algo)
## chm_95h_segmented_crowns = delineate_crowns(chm_95h_segmented, type = 'convex')
## chm_95h_segmented_crowns_spplot = sp::plot(chm_95h_segmented_crowns, cex=0.000001, axes = TRUE, alpha=0.1)

# Point Cloud Pipeline: Derive raster
## chm_95h_segmented_crowns_sf = st_as_sf(chm_95h_segmented_crowns, coords = c("XTOP", "YTOP"), st_crs=2193)
## psych::describe(chm_95h_segmented_crowns_sf$treeID)
## raster_template = rast(ext(chm_95height_raster), resolution = 1, crs = st_crs(chm_95height_raster)$wkt)
## segmented_crowns_rast = rasterize(vect(chm_95h_segmented_crowns_sf), raster_template, fun = sum, touches = TRUE)
## stem_count_ha_rast = segmented_crowns_rast*100 #convert from 1m resolution 100m-sq fpr stems per hectare
## stem_count_ha_raster = raster::raster(stem_count_ha_rast, filename = "~/Desktop/Summit_Forestry/stands/stem_count_ha_raster.tif")
## stem_count_ha_raster = writeRaster(stem_count_ha_raster, filename = "~/Desktop/Summit_Forestry/stands/stem_count_ha_raster.tif", overwrite=TRUE)
## stem_count_ha_raster = raster::raster(stem_count_ha_rast, filename = "~/Desktop/Summit_Forestry/stands/stem_count_ha_raster.tif")
## plot(segmented_crowns_rast)
## plot(stem_count_ha_raster)


```

# Outputs: Compare Stem Maps in Polygon, Heatmap and Histograms

```{r, fig.show='hold', out.width="50%", echo=FALSE}
stem_count_sf = st_as_sf(stem_count)
plot(st_geometry(las_tile_ahbau_ttops_sf["treeID"]), add=TRUE, cex = 0.001, pch=19, col = 'red', alpha=0.8)
plot(stem_count_raster, main="Quesnel Developed Blocks WSVHA", cex.main=0.8, maxpixels=22000000)
hist(stem_count_raster, main="Quesnel Developed Blocks WSVHA", cex.main=0.8, maxpixels=22000000) 
```

